---
title: 'Replication Code: Hospital Attacks Sudan'
author: "Lilli Blank, Franziska Gekeler"
date: "03/13/2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
  message = FALSE)
```

## Why do conflict parties attack hospitals in Sudan? 
This markdown shows the code used for our thesis "Why do conflict parties attack health care infrastructure? Understanding attacks on hospitals in Sudan". You'll find more information on the data sources in Section 3, for possible replication. The code is structured as follows:

### Table of contents

1. [Loading the necessary packages](#loading-the-necessary-packages)
2. [Load the datasets we use](#load-the-datasets-we-use)
3. [Create spatial unit of analysis: grids](#create-spatial-unit-of-analysis-grids)
4. [Prepare datasets](#prepare-datasets)
5. [Create spatial and temporal unit of analysis (grid week)](#create-spatial-and-temporal.unit-of-analysis-grid-week)
6. [Descriptive analysis](#descriptive-analysis)
7. [Merge all data (ZINB preparation)](#merge-all-data-zinb-preparation)
8. [ZINB analysis](#zinb-analysis)
9. [Statistical and robustness checks](#statistical-and-robustness-checks)

## Loading the necessary packages
```{r load-packages}
library(broom)
library(dplyr)
library(fuzzyjoin)
library(ggplot2)
library(leaflet)
library(lubridate)
library(readxl)
library(reshape2)
library(scales)
library(tidyverse)
library(stargazer)
library(sf)
library(sp)
library(terra) 
library(knitr)
library(pscl)
library(performance)
library(lmtest)
library(MASS)
```

## Load the datasets we use 
```{r load data}
ACLED_data <- read.csv("/Users/Lilli/Documents/R/MasterThesis/ACLED Data_2025-11-04.csv")
InsecurityInsight_data <- read_excel (("/Users/Lilli/Documents/R/MasterThesis/2016-2025-attacks-on-health-care-incident-data.xlsx"))
ConflictIntensity_data <- read.csv("/Users/Lilli/Documents/R/MasterThesis/ACLED Data_2025-11-04_ALL-ATTACKS.csv")
Geo_data1 <- st_read(("/Users/Lilli/Documents/R/MasterThesis/gadm41_SDN_0.shp")) 
pop_raster <- rast("/Users/Lilli/Documents/R/MasterThesis/sdn_pop_2023_CN_1km_R2025A_UA_v1.tif")

```

## Create spatial unit of analysis: grids 
```{r create grids}
#plot Sudan and create grids
#Plotting Sudan (as a whole, without grids)----
Sudan <- Geo_data1

print(Sudan)
plot(st_geometry(Sudan))

st_crs(Sudan)
Sudan <- st_set_crs(Sudan, 4326)
sudan_meters <- st_transform(Sudan, 3857)
st_bbox(sudan_meters)

#We're creating grids now---- 
grid_size <- 30000   # 30 km

grid <- st_make_grid(
  sudan_meters,
  cellsize = grid_size,
  square = TRUE
)

grid <- st_sf(
  grid_id = 1:length(grid),
  geometry = grid
)

grid_sudan <- st_intersection(grid, sudan_meters)

nrow(grid_sudan)

#create master-grid
grid_sf <- grid_sudan %>%
  mutate(grid_id = row_number())

#plotting of grids
spatial_unit_of_analysis <- ggplot() +
  geom_sf(data = grid_sudan, fill = NA, color = "grey70", size = 0.15) +
  geom_sf(data = sudan_meters, fill = NA, color = "black", size = 0.6) +
  theme_minimal(base_family = "Times New Roman") +
  labs(
    title = "30 × 30 km Grid Cells in Sudan",
    subtitle = "Spatial units of analysis",
    caption = "Source: Author's own grid based on GADM boundary"
  )

print(spatial_unit_of_analysis)
```

## Prepare datasets
```{r prepare data}
#Data Preparation -> descriptive Analysis

#Insecurity Insight /SHCC Data-Set ----
#Filter for Sudan only
sudan_HIA_data <- subset(InsecurityInsight_data, Country == "Sudan")
# put date into date format
sudan_HIA_data$event_date <- as.Date(sudan_HIA_data$Date, format = "%Y-%m-%d")
#filter events starting from 2023
sudan_HIA_data_2023 <- sudan_HIA_data %>%
  filter(event_date >= as.Date("2023-01-01"))


#filtering attacks on hospitals (health facilities) - Sudan_hospital_attacks_data is Data-Set we're working with
sudan_hospital_attacks_data <- sudan_HIA_data_2023 %>%
  filter(
    `Number of Attacks on Health Facilities Reporting Destruction` > 0 |
      `Number of Attacks on Health Facilities Reporting Damaged` > 0 |
      `Forceful Entry into Health Facility` > 0 |
      `Occupation of Health Facility` > 0
  )

#filter for events with Geo-precision
sudan_hospital_attacks_data <- sudan_hospital_attacks_data %>%
  dplyr::filter(!(`Geo Precision` == '(4) Province, State, Governorate'| `Geo Precision` == ' (6) Country' ))


#bring it in geometric form 
sudan_hospital_attacks_sf <- st_as_sf(
  sudan_hospital_attacks_data,
  coords = c("Longitude", "Latitude"),
  crs = 4326
)

sudan_hospital_attacks_sf <- st_transform(
  sudan_hospital_attacks_sf,
  st_crs(sudan_meters)
)

attacks_coords <- sudan_hospital_attacks_sf %>%
  dplyr::mutate(
    x = st_coordinates(.)[,1],
    y = st_coordinates(.)[,2]
  )

attacks_by_location <- attacks_coords %>%
  st_drop_geometry() %>%
  dplyr::count(x, y, name = "n_attacks")

attacks_locations_sf <- st_as_sf(
  attacks_by_location,
  coords = c("x", "y"),
  crs = st_crs(sudan_meters)
)

#prepare Insecurity Insight Data for grid week ---- 
  sudan_hospital_attacks_sf <- sudan_hospital_attacks_sf %>%
  mutate(
    event_date = as.Date(event_date),
    week = floor_date(event_date, "week")
  )

#map attacks on grids
#first bring data to common metric coordinate
sudan_hospital_attacks_sf <- st_transform(
  sudan_hospital_attacks_sf,
  st_crs(grid_sudan)
)

hospital_attacks_grid <- st_join(
  sudan_hospital_attacks_sf,
  grid_sudan,
  left = FALSE
)

hospital_attacks_grid_week <- hospital_attacks_grid %>%
  st_drop_geometry() %>%
  count(grid_id, week, name = "hospital_attacks")

#Filter for RSF and SAF - attacks_by_actor new dataset 
attacks_by_actor <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  dplyr::count(`Reported Perpetrator Name`, name = "n_attacks") %>%
  dplyr::arrange(desc(n_attacks))

#see how many attacks per actor 
attacks_by_actor

#see percentage of attacks per actor
attacks_by_actor <- attacks_by_actor %>%
  dplyr::mutate(
    share = round(n_attacks / sum(n_attacks) * 100, 1)
  )
attacks_by_actor

#ACLED territorial changes Data ----
territorial_change_data <- ACLED_data

#filter for events with geo and time precision lower than 3
territorial_change_data <- territorial_change_data %>%
  dplyr::filter(!(geo_precision == 3 | time_precision == 3))

#check, what kinds of sub_event are in the data
territorial_change_data %>% count(sub_event_type, sort =TRUE)
#filter for the territorial changes
territorial_events <- territorial_change_data %>%
  filter(sub_event_type %in% c(
    "Government regains territory",
    "Non-state actor overtakes territory", 
    "Non-violent transfer of territory"
  ))

territorial_events_sf <- territorial_events %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(st_crs(grid_sudan))

#ACLED Data: prepare data for conflict intensity ----
conflict_intensity_data <- ConflictIntensity_data

#check the disorder_type, and event_type
conflict_intensity_data %>% count(disorder_type, sort =TRUE)

conflict_intensity_data %>% count(event_type, sort =TRUE)

#filter for relevant events and mutate dates for further calculation
conflict_intensity_clean <- conflict_intensity_data %>%
  filter(disorder_type == "Political violence") %>%
  mutate(
    event_date = as.Date(event_date),
    week = floor_date(event_date, "week")
  ) %>%
  filter(event_date >= as.Date("2023-01-01"))

#sanity check
nrow(conflict_intensity_clean)
summary(conflict_intensity_clean$event_date)
table(conflict_intensity_clean$event_type)

#lose the events where Geo-precision and time precision has value 3
conflict_intensity_clean <- conflict_intensity_clean %>%
  dplyr::filter(!(geo_precision == 3 | time_precision == 3))

#check if entries with value 3 are gone 
sum(conflict_intensity_clean$geo_precision == 3, na.rm = TRUE)
sum(conflict_intensity_clean$time_precision == 3, na.rm = TRUE)

#mutate in sf 
conflict_intensity_sf <- conflict_intensity_clean %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(st_crs(grid_sudan))
#and map on grids
conflict_grid <- st_join(
  conflict_intensity_sf,
  grid_sudan,
  left = FALSE
)
#and aggregate conflict intensity per grid week 
conflict_intensity_grid_week <- conflict_grid %>%
  st_drop_geometry() %>%
  count(grid_id, week, name = "conflict_intensity")

summary(conflict_intensity_grid_week$conflict_intensity)

#prepare WorldPop Data for control variable ----

pop_raster
# 0) Check: grid_id darf NICHT neu nummeriert werden
stopifnot("grid_id" %in% names(grid_sudan))

# 2) Raster in CRS von grid_sudan projizieren 
pop_raster_proj <- project(pop_raster, st_crs(grid_sudan)$wkt)

# 3) grid_sudan -> SpatVector
grid_vect <- vect(grid_sudan)

# 4) Population pro Grid extrahieren (Summe)
population_grid <- extract(
  pop_raster_proj,
  grid_vect,
  fun = sum,
  na.rm = TRUE
)

#rename column
population_grid <- population_grid %>%
  rename(population_2023 = global_pop_2023_CN_1km_R2025A_v1)

names(population_grid)
```

## Create spatial and temporal unit of analysis (grid week)
```{r grid-week creation}
#Creation of grid-weeks 
#create vector of weeks 
weeks <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  mutate(week = floor_date(Date, "week")) %>%
  distinct(week)
#grid x week
panel_df <- expand.grid(
  grid_id = grid_sf$grid_id,
  week = weeks$week
)

#aggregate territorial events on grid-week

territorial_events_sf <- territorial_events_sf %>%
  mutate(
    event_date = as.Date(event_date),
    week = floor_date(event_date, "week")
  )

territorial_grid <- st_join(
  territorial_events_sf,
  grid_sudan,
  left = FALSE
)

territorial_counts <- territorial_grid %>%
  st_drop_geometry() %>%
  count(grid_id, week, name = "n_territorial_events")

#check (for sanity)
summary(territorial_counts$n_territorial_events)
sum(is.na(territorial_counts$week))
```

## Descriptive analysis 
```{r descriptive analysis}
#helper function ----
save_png <- function(plot, name, w = 18, h = 12,
                     path = "output") {
  dir.create(path, showWarnings = FALSE)
  ggsave(
    filename = file.path(path, paste0(name, ".png")),
    plot = plot,
    width = w, height = h, units = "cm",
    dpi = 300,
    device = "png"
  )
}

#Plot: counting the attacks against hospitals ----
weekly_attacks <- sudan_hospital_attacks_data %>%
  mutate(week = floor_date(event_date, "week")) %>%
  group_by(week) %>%
  summarise(count = n())

# create plot
p1 <- ggplot(weekly_attacks, aes(x = week, y = count)) +
  geom_line() +
  labs(title = "Attacks on hospitals in Sudan",
       x = "Week",
       y = "Number of attacks") +
  theme_minimal(base_family = "Times New Roman")
print(p1)
#check with conflict intensity
weekly_attacks_2 <- conflict_intensity_clean %>%
  mutate(week = floor_date(event_date, "week")) %>%
  group_by(week) %>%
  summarise(count = n())

scale_factor <- max(weekly_attacks$count, na.rm = TRUE) /
  max(weekly_attacks_2$count, na.rm = TRUE)

p2 <- ggplot() +
  geom_line(
    data = weekly_attacks,
    aes(x = week, y = count),
    linewidth = 1
  ) +
  geom_line(
    data = weekly_attacks_2,
    aes(x = week, y = count * scale_factor),
    linetype = "dashed",
    linewidth = 1
  ) +
  scale_y_continuous(
    name = "Number of hospital attacks",
    sec.axis = sec_axis(
      ~ . / scale_factor,
      name = "Conflict intensity"
    )
  ) +
  labs(
    title = "Hospital attacks and conflict intensity in Sudan",
    x = "Week"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(p2)

#define baseline and escalation ans (incl. rubustness check) ----
#rename column count
weekly_attacks <- weekly_attacks%>%
rename(n_attacks = count )

threshold <- quantile(weekly_attacks$n_attacks, 0.80, na.rm = TRUE)

weekly_attacks <- weekly_attacks %>%
  mutate(
    escalation_80 = ifelse(n_attacks >= threshold, "Escalation", "Baseline")
  )
#what's our 80th percentile
p80 <- quantile(weekly_attacks$n_attacks, probs = 0.80, na.rm = TRUE)
p80

#check distribution
summary(weekly_attacks$n_attacks)
#histogram 
hist(weekly_attacks$n_attacks, breaks = 20)

attack_threshold <- 4
#plot weekly series, threshold and the escalation marking
escalation_plot_80 <- ggplot(weekly_attacks, aes(x = week, y = n_attacks)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = attack_threshold, linetype = "dashed", linewidth = 0.5) +
  geom_point(aes(color = escalation_80), size = 2) +
  scale_color_manual(values = c("Baseline" = "grey", "Escalation" = "red")) +
  labs(
    x = "Week",
    y = "Number of attacks on hospitals",
    title = "Weekly attacks on hospitals in Sudan",
    subtitle = "Escalation defined as weeks with ≥ 4 attacks (≈ 80th percentile)",
    color = ""
  ) +
  theme_minimal(base_family = "Times New Roman")

print(escalation_plot_80)

#robustness checks by plotting the percentage of escalation weeks ----
#plot cumulated attacks ----
cumulated_attacks <- ggplot(weekly_attacks, aes(x = n_attacks)) +
  stat_ecdf(geom = "step") +
  geom_hline(yintercept = 0.8, linetype = "dashed") +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 1)
  ) +
  labs(
    title = "Cumulative distribution of weekly hospital attacks",
    x = "Number of hospital attacks per week",
    y = "Cumulative share of weeks"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(cumulated_attacks)

cumulated_attacks_zoomed <- ggplot(data = weekly_attacks, aes(x = n_attacks)) +
  stat_ecdf(geom = "step") +
  coord_cartesian(xlim = c(0, 10)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_x_continuous(breaks = 0:10) +
  labs(
    title = "Cumulative distribution of weekly hospital attacks (zoomed)",
    x = "Number of hospital attacks per week",
    y = "Cumulative share of weeks"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(cumulated_attacks_zoomed)

#Plotting InsecurityInsight Data -- put attacks on hospitals on map and show number of attacks on the hospitals ----

#plot repeated attacks on hospital on map (data: attacks_locations_sf)

#setting key places for map
key_places <- data.frame(
  name = c("Khartoum", "al-Fashir", "Nyala", "El Geneina"),
  lon = c(32.5599, 25.3500, 24.8800, 22.4400),
  lat = c(15.5007, 13.6300, 12.0500, 13.4500)
)

key_places_sf <- st_as_sf(key_places,
                          coords = c("lon", "lat"),
                          crs = 4326)

# project to same CRS as your map
key_places_sf <- st_transform(key_places_sf, st_crs(sudan_meters))

repeated_attacks_plot <- ggplot() +
  
  geom_sf(data = sudan_meters, fill = NA, color = "black", size = 0.6) +
  
  geom_sf(
    data = attacks_locations_sf,
    aes(color = n_attacks),
    size = 1.3,
    alpha = 0.85
  ) +
  
  # Key place markers (white circle with black border)
  geom_sf(
    data = key_places_sf,
    shape = 21,
    fill = NA,
    color = "black",
    size = 1.0,
    stroke = 0.5
  ) +
  
  # Labels for key places
  geom_sf_text(
    data = key_places_sf,
    aes(label = name),
    size = 3,
    nudge_y = 60000,
    family = "Times New Roman"
  ) +
  
  scale_color_viridis_c(
    option = "rocket",
    direction = -1,
    name = "Number of attacks"
  ) +
  
  theme_minimal(base_family = "Times New Roman") +
  
  labs(
    title = "Repeated attacks on hospitals in Sudan",
    subtitle = "Colors indicate number of attacks at the same location",
    caption = "Yellow: at least 1 attack, Purple: over 40 attacks; Data source: Insecurity Insight"
  )

print(repeated_attacks_plot)

#no plot but table (see how often several hospitals have been attacked)

attacks_per_location <- attacks_coords %>%
  st_drop_geometry() %>%
  dplyr::count(x, y, name = "n_attacks") %>%
  dplyr::arrange(desc(n_attacks))

head(attacks_per_location, 25)

#Daily attacks on hospitals by RSF and SAF ----
# plot
sudan_hospital_attacks_perp <- sudan_hospital_attacks_data %>%
  filter(`Reported Perpetrator Name` %in% c("Rapid Support Forces", "Sudanese Armed Forces"))

daily_counts <- sudan_hospital_attacks_perp %>%
  count(event_date, `Reported Perpetrator Name`)


p5 <- ggplot(daily_counts,
       aes(x = event_date, 
           y = n, 
           color = `Reported Perpetrator Name`)) +
  geom_line(size = 1) +
  labs(
    title = "Daily Attacks on hospitals: RSF vs SAF",
    x = "Date",
    y = "Number of Attacks",
    scale_color_manual(
      name = "Perpetrator",
      values = c(
        "SAF" = "#B22222",
        "RSF" = "#1F4E79",
        "RSF + SAF" = "#7A1FA2" )
  )) +
  theme_minimal(base_family = "Times New Roman")

print(p5)

#Who attacks where and what?
#filter for reported perpetrator and geometry
sudan_hospital_attacks_actor_sf <- sudan_hospital_attacks_sf %>%
  dplyr::select(
    actor = `Reported Perpetrator Name`,
    geometry
  )

#exclude NAs and NIs
sudan_hospital_attacks_actor_sf_filtered <- sudan_hospital_attacks_actor_sf %>%
  mutate(
    actor_clean = case_when(
      actor == "Sudanese Armed Forces" ~ "SAF",
      actor == "Rapid Support Forces" ~ "RSF",
      actor %in% c(
        "Rapid Support Forces, Sudanese Armed Forces",
        "Sudanese Armed Forces, Rapid Support Forces"
      ) ~ "RSF + SAF",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(actor_clean))

# plot with key places and attacks per actor 

actor_attacks_plot_color_new <- ggplot() +
  
  geom_sf(
    data = sudan_meters,
    fill = NA,
    color = "black",
    size = 0.6
  ) +
  
  geom_sf(
    data = sudan_hospital_attacks_actor_sf_filtered,
    aes(color = actor_clean),
    size = 1.3,
    alpha = 0.85
  ) +
  
  # Key place markers (transparent circle with black border)
  geom_sf(
    data = key_places_sf,
    shape = 21,
    fill = NA,
    color = "black",
    size = 1,
    stroke = 0.5
  ) +
  
  # Labels for key places
  geom_sf_text(
    data = key_places_sf,
    aes(label = name),
    size = 3,
    nudge_y = 60000,
    family = "Times New Roman"
  ) +
  
  scale_color_manual(
    name = "Reported perpetrator",
    values = c(
      "SAF" = "#B22222",
      "RSF" = "#1F4E79",
      "RSF + SAF" = "#7A1FA2"
    ),
    breaks = c("SAF", "RSF", "RSF + SAF")
  ) +
  
  theme_minimal(base_family = "Times New Roman") +
  
  labs(
    title = "Attacks on hospitals in Sudan by conflict party",
    subtitle = "Red: SAF, Blue: RSF, Purple: jointly reported (RSF + SAF)",
    caption = "Data source: Insecurity Insight"
  )

print(actor_attacks_plot_color_new)

#with grey dots for no NIs
actor_attacks_plot_color <- ggplot() +
  geom_sf(
    data = sudan_meters,
    fill = NA,
    color = "black",
    size = 0.6
  ) +
  geom_sf(
    data = sudan_hospital_attacks_actor_sf,
    aes(color = actor),
    size = 1.3,
    alpha = 0.85
  ) +
  scale_color_manual(
    name = "Perpetrator",
    values = c(
      "Sudanese Armed Forces" = "#C00000",  
      "Rapid Support Forces" = "#1F4E79"    
    )
  ) +
  theme_minimal(base_family = "Times New Roman") +
  labs(
    title = "Attacks on hospitals in Sudan by conflict party",
    subtitle = "Red: SAF, Blue: RSF",
    caption = "Data source: Insecurity Insight"
  )

print(actor_attacks_plot_color)

# Weapons and actors ----
#which weapons are used? ----
weapons_overall <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  dplyr::count(`Weapon Carried/Used`, name = "n_attacks") %>%
  dplyr::arrange(desc(n_attacks))

#in percentage
weapons_overall <- weapons_overall %>%
  dplyr::mutate(
    share = round(n_attacks / sum(n_attacks) * 100, 1)
  )
weapons_overall

#group weapons, to see which actor uses which weapons
weapons_grouped <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  dplyr::mutate(
    weapon_group = dplyr::case_when(
      
      `Weapon Carried/Used` %in% c(
        "Shelling", "Artillery", "Mortar", "Rocket",
        "Missile", "Unspecified Explosive"
      ) ~ "Explosives / Heavy fire",
      
      `Weapon Carried/Used` %in% c(
        "Aerial Bomb: Plane", "Aerial Bomb: Drone",
        "Aerial Bomb: Missile", "Aerial Bomb: Plane, Artillery"
      ) ~ "Air-delivered weapons",
      
      `Weapon Carried/Used` == "Firearms" ~ "Firearms",
      
      `Weapon Carried/Used` == "Arson" ~ "Arson",
      
      `Weapon Carried/Used` == "Fist and Foot" ~ "Physical force",
      
      TRUE ~ "Unknown"
    )
  ) %>%
  dplyr::count(weapon_group, name = "n_attacks") %>%
  dplyr::mutate(
    share = round(n_attacks / sum(n_attacks) * 100, 1)
  ) %>%
  dplyr::arrange(desc(n_attacks))

weapons_grouped

#Add type of actor (RSF & SAF x type of weapon)----
attacks_actor_weapon <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  dplyr::mutate(
    
    actor_group = dplyr::case_when(
      grepl("Rapid Support Forces|RSF", `Reported Perpetrator Name`, ignore.case = TRUE) ~ "RSF",
      grepl("Sudanese Armed Forces|SAF|Sudan Armed Forces", `Reported Perpetrator Name`, ignore.case = TRUE) ~ "SAF",
      grepl("Police|Security", `Reported Perpetrator Name`, ignore.case = TRUE) ~ "State (other)",
      grepl("Unknown|Unidentified", `Reported Perpetrator Name`, ignore.case = TRUE) ~ "Unknown",
      TRUE ~ "Other armed actors"
    ),
    
    weapon_group = dplyr::case_when(
      `Weapon Carried/Used` %in% c(
        "Shelling", "Artillery", "Mortar", "Rocket",
        "Missile", "Unspecified Explosive"
      ) ~ "Explosives / Heavy fire",
      
      `Weapon Carried/Used` %in% c(
        "Aerial Bomb: Plane", "Aerial Bomb: Drone",
        "Aerial Bomb: Missile", "Aerial Bomb: Plane, Artillery"
      ) ~ "Air-delivered weapons",
      
      `Weapon Carried/Used` == "Firearms" ~ "Firearms",
      `Weapon Carried/Used` == "Arson" ~ "Arson",
      `Weapon Carried/Used` == "Fist and Foot" ~ "Physical force",
      
      TRUE ~ "Unknown"
    )
  )

actor_weapon_counts <- attacks_actor_weapon %>%
  dplyr::count(actor_group, weapon_group, name = "n_attacks") %>%
  dplyr::arrange(desc(n_attacks))

#Compare RSF and SAF, wepaon shares, incl. heatmap
actor_weapon_shares <- actor_weapon_counts %>%
  dplyr::group_by(actor_group) %>%
  dplyr::mutate(
    share_actor = round(n_attacks / sum(n_attacks) * 100, 1)
  ) %>%
  dplyr::ungroup()

actor_weapon_shares

#heatmap for who x what
heatmap_weapon_shares <- ggplot(actor_weapon_shares,
       aes(x = weapon_group, y = actor_group, fill = share_actor)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "rocket",
    direction = -1,
    name = "Share of attacks (%)"
  ) +
  labs(
    x = "Weapon category",
    y = "Actor",
    title = "Weapons used in attacks on hospitals by actor"
  ) +
  theme_minimal(base_family = "Times New Roman") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(heatmap_weapon_shares)

#aggregate attacks per week 
attacks_weekly <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  mutate(
    week = floor_date(Date, unit = "week", week_start = 1)
  ) %>%
  count(week, name = "n_attacks")

#Show: actors and attacks in plot ----
sudan_hospital_attacks_sf %>% st_drop_geometry () %>% count (`Reported Perpetrator Name`, sort = TRUE)

attacks_weekly_actor <- sudan_hospital_attacks_sf %>%
  st_drop_geometry() %>%
  mutate(
    week = floor_date(Date, unit = "week", week_start = 1),
    actor_group = case_when(
      grepl("Rapid Support Forces",`Reported Perpetrator Name` , ignore.case = TRUE) ~ "RSF",
      grepl("Sudanese Armed Forces", `Reported Perpetrator Name`, ignore.case = TRUE) ~ "SAF",
      TRUE ~ "Other"
    )
  ) %>%
  count(week, actor_group)
#plot it
plot3 <- ggplot(
  attacks_weekly_actor,
  aes(x = week, y = n, color = actor_group)
) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(
    values = c(
      "RSF"   = "#7B1E1E",
      "SAF"   = "#1F4E79",  
      "Other" = "grey60"  
    ),
    breaks = c("RSF", "SAF", "Other")
  ) +
  labs(
    x = "Week",
    y = "Number of attacks",
    color = "Actor"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(plot3)
```

## Merge all data (ZINB preparation)
```{r merging all data}
#merge all data
#do we have all we need? check:

#all grid-week combinations: 
# panel_df
#all hospital attacks per grid week:
# hospital_attacks_grid_week
#all territorial counts:
# territorial_counts
#conflict intensity per grid-week:
# conflict_intensity_grid_week
#population per grid:
# population_grid

#make sure to have the same time period of the data
# Check min / max week in each Data-Set
range(hospital_attacks_grid_week$week, na.rm = TRUE)
range(territorial_counts$week, na.rm = TRUE)
range(conflict_intensity_grid_week$week, na.rm = TRUE)

start_week <- as.Date("2023-03-26")  # latest common start
end_week   <- as.Date("2024-11-03")  # earliest common end

hospital_attacks_grid_week <- hospital_attacks_grid_week %>%
  dplyr::filter(week >= start_week, week <= end_week)

territorial_counts <- territorial_counts %>%
  dplyr::filter(week >= start_week, week <= end_week)

conflict_intensity_grid_week <- conflict_intensity_grid_week %>%
  dplyr::filter(week >= start_week, week <= end_week)

#check if correctly filtered
range(hospital_attacks_grid_week$week)
range(territorial_counts$week)
range(conflict_intensity_grid_week$week)
#check numbers of weeks per set
Reduce(
  intersect,
  list(
    unique(hospital_attacks_grid_week$week),
    unique(territorial_counts$week),
    unique(conflict_intensity_grid_week$week)
  )
) |> length()
#correct the weeeks of territorial counts
all_weeks <- seq(
  from = as.Date("2023-03-26"),
  to   = as.Date("2024-11-03"),
  by   = "week"
)

territorial_counts <- territorial_counts %>%
  tidyr::complete(
    grid_id,
    week = all_weeks,
    fill = list(n_territorial_events = 0)
  )

range(hospital_attacks_grid_week$week)
range(territorial_counts$week)
range(conflict_intensity_grid_week$week)

#check: same keys? 
lapply(
  list(
    panel_df,
    hospital_attacks_grid_week,
    territorial_counts,
    conflict_intensity_grid_week
  ),
  function(x) names(x)
)

#merge (left joins, no inner join or we lose structural zeros)
panel_master <- panel_df %>%
  left_join(
    hospital_attacks_grid_week,
    by = c("grid_id", "week")
  ) %>%
  left_join(
    territorial_counts,
    by = c("grid_id", "week")
  ) %>%
  left_join(
    conflict_intensity_grid_week,
    by = c("grid_id", "week")
  )
#transform NAs to 0, for ZINB essential
panel_master <- panel_master %>%
  mutate(
    hospital_attacks = ifelse(is.na(hospital_attacks), 0, hospital_attacks),
    n_territorial_events = ifelse(is.na(n_territorial_events), 0, n_territorial_events),
    conflict_intensity = ifelse(is.na(conflict_intensity), 0, conflict_intensity)
  )
#sanity and mandatory checks
summary(panel_master$hospital_attacks)
mean(panel_master$hospital_attacks)
#mean is almost zero: 0.0005526495
#check logic
panel_master %>%
  dplyr::filter(hospital_attacks > 0) %>%
  dplyr::select(grid_id, week, hospital_attacks, conflict_intensity, n_territorial_events) %>%
  head(10)

#check wether it happens often that hosptial attack = 1 and rest =0
hospital_only <- panel_master %>%
  dplyr::filter(
    hospital_attacks > 0,
    conflict_intensity == 0,
    n_territorial_events == 0
  ) %>%
  dplyr::select(grid_id, week, hospital_attacks, conflict_intensity, n_territorial_events)

nrow(hospital_only)
#share of hospital attack grid-weeks (without something else happening)
total_hospital_weeks <- panel_master %>%
  filter(hospital_attacks > 0) %>%
  nrow()

nrow(hospital_only) / total_hospital_weeks

#share of hospital attacks with conflict in grid-week
hospital_with_conflict <- panel_master %>%
  filter(
    hospital_attacks > 0,
    conflict_intensity > 0
  )
nrow(hospital_with_conflict)
nrow(hospital_only)

#share of hospital attacks and territorial changes in grid-week
hospital_with_territorial <- panel_master %>%
  filter(
    hospital_attacks > 0,
    n_territorial_events > 0
  )

nrow(hospital_with_territorial)
nrow(hospital_only)

#and share of both 
hospital_with_territorial_conflict <- panel_master %>%
  filter(
    hospital_attacks > 0,
    n_territorial_events > 0, 
    conflict_intensity > 0
  )

nrow(hospital_with_territorial_conflict)

#testing things to check whether merge was correct
sum(conflict_intensity_grid_week$conflict_intensity)
nrow(conflict_intensity_clean)
nrow(conflict_intensity_grid_week)

#creating lags - because attacks do not (or do??) occur time-delayed
panel_master_lag <- panel_master %>%
  arrange(grid_id, week) %>%
  group_by(grid_id) %>%
  mutate(
    conflict_intensity_lag1 = lag(conflict_intensity, 1),
    territorial_events_lag1 = lag(n_territorial_events, 1)
  ) %>%
  ungroup()
#sanity check
summary(panel_master_lag$conflict_intensity_lag1)
summary(panel_master_lag$territorial_events_lag1)

#check share: Lag share - Did something happen before (in grid), that might have influenced the attack? 
panel_master_lag %>%
  filter(hospital_attacks > 0) %>%
  summarise(
    share_conflict_lag1 = mean(conflict_intensity_lag1 > 0, na.rm = TRUE),
    share_territorial_lag1 = mean(territorial_events_lag1 > 0, na.rm = TRUE)
  )

#check share without lag: Contemporaneous Share - Does something happen in the same grid-week?
panel_master %>%
  filter(hospital_attacks > 0) %>%
  summarise(
    share_conflict_now = mean(conflict_intensity > 0),
    share_territorial_now = mean(n_territorial_events > 0)
  )

#visualization of repeated attacks in one grid 
# point at the repeated-attack coordinate
pt47 <- st_as_sf(
  data.frame(x = 3617883, y = 1746892),
  coords = c("x","y"),
  crs = st_crs(sudan_meters)   # same metric CRS
)

# get the grid cell (and its id) that contains the point
grid47 <- st_join(pt47, grid["grid_id"], join = st_within)

grid_id_47 <- grid47$grid_id[1]
grid_id_47


grid404_plot <- panel_master_lag %>%
  filter(grid_id == grid_id_47) %>%
  ggplot(aes(x = week, y = hospital_attacks)) +
  geom_line() +
  labs(
    title = "Weekly hospital attacks in the grid containing the most repeated coordinates",
    subtitle = paste0("Grid: ", grid_id_47, " | Location: (3617883, 1746892)"),
    x = "Week",
    y = "Number of hospital attacks"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(grid404_plot)
#merge population grid on panel_master_lag ----
population_grid_clean <- population_grid %>%
  rename(grid_id = ID)

panel_master_lag <- panel_master_lag %>%
  left_join(population_grid_clean, by = "grid_id")

sum(is.na(panel_master_lag$population_2023))

#loose grids where population = 0 
panel_master_lag <- panel_master_lag %>%
  mutate(
    population_2023 = as.numeric(unlist(population_2023))
  )

panel_master_lag <- panel_master_lag %>%
  mutate(
    population = ifelse(is.na(population_2023) | is.nan(population_2023), 0, population_2023)
  )

panel_master_lag <- panel_master_lag %>%
  filter(population > 0)

#check
str(panel_master_lag$population)
table(panel_master_lag$population == 0)

panel_master_lag <- panel_master_lag %>%
  mutate(log_population = log(population))

```

## ZINB analysis
```{r ZINB models}
#Zero-Inflated Negative Binomial Regression 

#ZINB_pop ----
zinb_pop <- zeroinfl(
  hospital_attacks ~ 
    territorial_events_lag1 +
    conflict_intensity + 
    n_territorial_events +
    log_population |
    conflict_intensity +
    log_population + 
    territorial_events_lag1 +
    n_territorial_events,
  data = panel_master_lag,
  dist = "negbin"
)

summary(zinb_pop)

zinb_pop_lag <- zeroinfl(
  hospital_attacks ~ 
    conflict_intensity_lag1 +
    territorial_events_lag1 +
    n_territorial_events +
    log_population |
    conflict_intensity_lag1 +
    log_population + 
    territorial_events_lag1 +
    n_territorial_events,
  data = panel_master_lag,
  dist = "negbin"
)

summary(zinb_pop_lag)

#actor specific ZINBs ----
#first: add actor specifications in panel_master_lag ----
hospital_attacks_actor <- sudan_hospital_attacks_sf %>%
  mutate(
    RSF = grepl("Rapid Support Forces", `Reported Perpetrator Name`, ignore.case = TRUE),
    SAF = grepl("Sudanese Armed Forces", `Reported Perpetrator Name`, ignore.case = TRUE)
  )

hospital_attacks_actor <- st_join(
  hospital_attacks_actor,
  grid_sudan %>% dplyr::select(grid_id),
  join = st_intersects
)
sum(is.na(hospital_attacks_actor$grid_id))

#aggregation grid-week-actor
attacks_gridweek <- hospital_attacks_actor %>%
  st_drop_geometry() %>%
  group_by(grid_id, week) %>%
  summarise(
    hospital_attacks_RSF = sum(RSF, na.rm = TRUE),
    hospital_attacks_SAF = sum(SAF, na.rm = TRUE),
    .groups = "drop"
  )

panel_master_lag_actors <- panel_master_lag %>%
  left_join(attacks_gridweek, by = c("grid_id", "week")) %>%
  mutate(
    hospital_attacks_RSF = ifelse(is.na(hospital_attacks_RSF), 0, hospital_attacks_RSF),
    hospital_attacks_SAF = ifelse(is.na(hospital_attacks_SAF), 0, hospital_attacks_SAF)
  )

#second: add territorial losses 
territorial_losses <- territorial_events_sf %>%
  mutate(
    loss_SAF = sub_event_type == "Non-state actor overtakes territory",
    loss_RSF = sub_event_type == "Government regains territory",
    loss_nonviolent = sub_event_type == "Non-violent transfer of territory"
  )
#put to grids
territorial_losses <- st_join(
  territorial_losses,
  grid_sudan %>% dplyr::select(grid_id),
  join = st_intersects
)
#aggregate on grid-week
territorial_losses_gridweek <- territorial_losses %>%
  st_drop_geometry() %>%
  group_by(grid_id, week) %>%
  summarise(
    territorial_losses_RSF = sum(loss_RSF, na.rm = TRUE),
    territorial_losses_SAF = sum(loss_SAF, na.rm = TRUE),
    territorial_transfers_nonviolent = sum(loss_nonviolent, na.rm = TRUE),
    .groups = "drop"
  )
#merge in panel
panel_master_lag_actors <- panel_master_lag_actors %>%
  left_join(territorial_losses_gridweek, by = c("grid_id", "week")) %>%
  mutate(
    territorial_losses_RSF = ifelse(is.na(territorial_losses_RSF), 0, territorial_losses_RSF),
    territorial_losses_SAF = ifelse(is.na(territorial_losses_SAF), 0, territorial_losses_SAF),
    territorial_transfers_nonviolent = ifelse(is.na(territorial_transfers_nonviolent), 0,
                                              territorial_transfers_nonviolent)
  )

#create lags for territorial losses
panel_master_lag_actors <- panel_master_lag_actors %>%
  arrange(grid_id, week) %>%
  group_by(grid_id) %>%
  mutate(
    territorial_losses_RSF_lag1 = lag(territorial_losses_RSF, 1),
    territorial_losses_SAF_lag1 = lag(territorial_losses_SAF, 1)
  ) %>%
  ungroup()
#sanity check
summary(panel_master_lag_actors$territorial_losses_RSF_lag1)
summary(panel_master_lag_actors$territorial_losses_SAF_lag1)

#check share: Lag share - Did something happen before (in grid), that might have influenced the attack? 
panel_master_lag_actors %>%
  filter(hospital_attacks_RSF > 0) %>%
  summarise(
    share_conflict_lag1 = mean(conflict_intensity_lag1 > 0, na.rm = TRUE),
    share_territorial_lag1 = mean(territorial_losses_RSF_lag1 > 0, na.rm = TRUE)
  )

#ZINB_RSF ----

#zinb_RSF - contemporaneous
# RSF – contemporaneous
zinb_RSF_t <- zeroinfl(
  hospital_attacks_RSF ~
    conflict_intensity +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population |
    conflict_intensity +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population,
  data = panel_master_lag_actors,
  dist = "negbin"
)

summary(zinb_RSF_t)
# RSF – lagged
zinb_RSF_lag <- zeroinfl(
  hospital_attacks_RSF ~
    conflict_intensity_lag1 +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population |
    conflict_intensity_lag1 +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population,
  data = panel_master_lag_actors,
  dist = "negbin"
)

summary(zinb_RSF_lag)

#model fit? 
AIC(zinb_RSF_t, zinb_RSF_lag)

#ZINB_SAF ---- 
#zinb SAF contemporaneous 
zinb_SAF_t <- zeroinfl(
  hospital_attacks_SAF ~
    conflict_intensity +
    territorial_losses_SAF +
    log_population |
    conflict_intensity +
    territorial_losses_SAF +
    log_population,
  data = panel_master_lag_actors,
  dist = "negbin"
)

summary(zinb_SAF_t)
# SAF – lagged
zinb_SAF_lag <- zeroinfl(
  hospital_attacks_SAF ~
    conflict_intensity_lag1 +
    territorial_losses_SAF +
    log_population |
    conflict_intensity_lag1 +
    territorial_losses_SAF +
    log_population,
  data = panel_master_lag_actors,
  dist = "negbin"
)

summary(zinb_SAF_lag)

#plotting the result of zinb


p_any_attack_zeroinfl <- function(model, newdata) {
  # linear predictors
  eta_zero  <- predict(model, newdata = newdata, type = "zero")
  eta_count <- predict(model, newdata = newdata, type = "count")
  
  # components
  pi <- plogis(eta_zero)        # structural-zero probability
  mu <- exp(eta_count)          # count mean
  theta <- model$theta          # NB dispersion
  
  # NB zero probability
  p_nb0 <- (theta / (theta + mu))^theta
  
  # overall Pr(Y >= 1)
  1 - (pi + (1 - pi) * p_nb0)
}


#prediction grids (scale identical for SAF & RSF)
make_grid_ci <- function(data, var, n = 100) {
  base <- data %>%
    summarise(across(where(is.numeric), ~ median(.x, na.rm = TRUE))) %>%
    slice(1)
  
  grid <- base[rep(1, n), , drop = FALSE]
  grid[[var]] <- seq(
    min(data[[var]], na.rm = TRUE),
    max(data[[var]], na.rm = TRUE),
    length.out = n
  )
  grid
}

#predicted probability plot RSF
grid_RSF <- make_grid_ci(panel_master_lag_actors, "conflict_intensity")

grid_RSF$pr_any <- p_any_attack_zeroinfl(zinb_RSF_t, grid_RSF)

ggplot(grid_RSF, aes(x = conflict_intensity, y = pr_any)) +
  
  geom_line(size = 0.5) +
  labs(
    x = "Conflict intensity",
    y = "Pr(≥1 hospital attack)",
    title = "RSF: Predicted probability of hospital attacks"
  ) +
  theme_minimal(base_family = "Times New Roman")

#predicted probability plot SAF
grid_SAF <- make_grid_ci(panel_master_lag_actors, "conflict_intensity")

grid_SAF$pr_any <- p_any_attack_zeroinfl(zinb_SAF_t, grid_SAF)

ggplot(grid_SAF, aes(x = conflict_intensity, y = pr_any)) +
  geom_line(size = 0.5) +
  labs(
    x = "Conflict intensity",
    y = "Pr(≥1 hospital attack)",
    title = "SAF: Predicted probability of hospital attacks"
  ) +
  theme_minimal(base_family = "Times New Roman")

#figure: SAF vs RSF 

grid_RSF$actor <- "RSF"
grid_SAF$actor <- "SAF"

plot_df <- bind_rows(
  grid_RSF %>% dplyr::select(conflict_intensity, pr_any, actor),
  grid_SAF %>% dplyr::select(conflict_intensity, pr_any, actor)
)

ggplot(plot_df,
       aes(x = conflict_intensity, y = pr_any, linetype = actor)) +
  geom_line(size = 0.5) +
  labs(
    x = "Conflict intensity",
    y = "Pr(≥1 hospital attack)",
    title = "Predicted probability of hospital attacks by actor",
    caption = "Predictions based on zero-inflated negative binomial models."
  ) +
  theme_minimal(base_family = "Times New Roman")

#plot RSF x territorial loss
grid_RSF_loss <- expand.grid(
  conflict_intensity = seq(
    min(panel_master_lag_actors$conflict_intensity, na.rm = TRUE),
    max(panel_master_lag_actors$conflict_intensity, na.rm = TRUE),
    length.out = 80
  ),
  territorial_losses_RSF = c(0, 1)
)

grid_RSF_loss$territorial_losses_RSF_lag1 <- median(
  panel_master_lag_actors$territorial_losses_RSF_lag1, na.rm = TRUE
)
grid_RSF_loss$log_population <- median(
  panel_master_lag_actors$log_population, na.rm = TRUE
)

grid_RSF_loss$pr_any <- p_any_attack_zeroinfl(zinb_RSF_t, grid_RSF_loss)

grid_RSF_loss$loss <- factor(
  grid_RSF_loss$territorial_losses_RSF,
  labels = c("No territorial loss", "Territorial loss")
)

ggplot(grid_RSF_loss,
       aes(x = conflict_intensity, y = pr_any, linetype = loss)) +
  geom_line(size = 0.5) +
  labs(
    x = "Conflict intensity",
    y = "Pr(≥1 hospital attack)",
    title = "RSF: Conflict intensity × territorial losses"
  ) +
  theme_minimal(base_family = "Times New Roman")

##
#structural zero probability vs conflict intensity
grid_RSF$pi <- plogis(predict(zinb_RSF_t, newdata = grid_RSF, type = "zero"))
grid_RSF$mu <- exp(predict(zinb_RSF_t, newdata = grid_RSF, type = "count"))

ggplot(grid_RSF, aes(x = conflict_intensity, y = pi)) +
  geom_line() +
  labs(x = "Conflict intensity", y = "Structural zero probability (π)",
       title = "RSF: Structural-zero probability vs conflict intensity") +
  theme_minimal(base_family = "Times New Roman")

#expected count (given at risk) vs conflict intensity 
ggplot(grid_RSF, aes(x = conflict_intensity, y = mu)) +
  geom_line() +
  labs(x = "Conflict intensity", y = "Count mean (μ)",
       title = "RSF: Expected count (given at risk) vs conflict intensity") +
  theme_minimal(base_family = "Times New Roman")

#test further (Plot:Zero inflation coefficients: SAF vs RSF) ----
summary(zinb_RSF_t)$coefficients
extract_zeroinfl_coefs <- function(model, component = c("count", "zero")) {
  component <- match.arg(component)
  
  coefs <- summary(model)$coefficients[[component]]
  
  data.frame(
    term = rownames(coefs),
    estimate = coefs[, "Estimate"],
    std.error = coefs[, "Std. Error"],
    z = coefs[, "z value"],
    p.value = coefs[, "Pr(>|z|)"],
    row.names = NULL
  )
}

#coefficient plots - SAF vs. RSF, zero part only 
zi_SAF <- extract_zeroinfl_coefs(zinb_SAF_t, "zero") %>%
  mutate(actor = "SAF")

zi_RSF <- extract_zeroinfl_coefs(zinb_RSF_t, "zero") %>%
  mutate(actor = "RSF")

zi_all <- bind_rows(zi_SAF, zi_RSF) %>%
  filter(term %in% c(
    "conflict_intensity",
    "conflict_intensity_lag1",
    "territorial_losses_SAF",
    "territorial_losses_RSF"
  )) %>%
  mutate(
    conf.low  = estimate - 1.96 * std.error,
    conf.high = estimate + 1.96 * std.error
  )

ggplot(zi_all, aes(x = estimate, y = term, shape = actor)) +
  geom_point(size = 2) +
  geom_errorbarh(
    aes(xmin = conf.low, xmax = conf.high),
    height = 0.2
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Effect on probability of being structurally attack-free (logit scale)",
    y = "",
    title = "Zero-inflation coefficients: SAF vs RSF",
    caption = "Negative values indicate a higher risk of hospital attacks."
  ) +
  theme_minimal(base_family = "Times New Roman")
```

## Statistical and robustness checks
```{r statistical checks}
## robustness and statistical checks 

## multicollinearity ----

#collinearity check for ZINB base model ----
vif_res <- check_collinearity(
  glm(
    hospital_attacks ~ territorial_events_lag1 + 
      n_territorial_events + 
      conflict_intensity +
      log_population,
    data = panel_master_lag,
    family = poisson()
  )
)

vif_df <- as.data.frame(vif_res)

names(vif_df)
vif_df
vif_df <- vif_df %>% rename(Parameter = Term)
vif_df
#plot for multi collinearity 
multicollinearity_plot <- ggplot(vif_df, aes(x = reorder(Parameter, VIF), y = VIF)) +
  geom_point(size = 3, color = "#1F4E79") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 10, linetype = "dotted", color = "grey50") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Variance Inflation Factor (VIF)",
    title = "Assessment of multicollinearity",
    subtitle = "All VIF values are well below conventional thresholds"
  ) +
  theme_minimal(base_family = "Times New Roman")

print(multicollinearity_plot)

ggsave(
  filename = "multicollinearity_plot.png",
  plot = multicollinearity_plot,    
  width = 18,
  height = 12,
  units = "cm", dpi = 300
)

#collinearity check for ZINB RSF model  -----
vif_RSF <- check_collinearity(
  glm(
    hospital_attacks_RSF ~
      conflict_intensity +
      territorial_losses_RSF +
      territorial_losses_RSF_lag1 +
      log_population,
    data = panel_master_lag_actors,
    family = poisson()
  )
)

vif_RSF_df <- as.data.frame(vif_RSF) %>%
  rename(Parameter = Term)

ggplot(vif_RSF_df, aes(x = reorder(Parameter, VIF), y = VIF)) +
  geom_point(size = 3, color = "#8B0000") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 10, linetype = "dotted", color = "grey50") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Variance Inflation Factor (VIF)",
    title = "Assessment of multicollinearity (RSF model)",
    subtitle = "No evidence of problematic multicollinearity"
  ) +
  theme_minimal(base_family = "Times New Roman")

#collinearity check for ZINB SAF model  -----
vif_SAF <- check_collinearity(
  glm(
    hospital_attacks_SAF ~
      conflict_intensity +
      territorial_losses_SAF +
      log_population,
    data = panel_master_lag_actors,
    family = poisson()
  )
)

vif_SAF_df <- as.data.frame(vif_SAF) %>%
  rename(Parameter = Term)

ggplot(vif_SAF_df, aes(x = reorder(Parameter, VIF), y = VIF)) +
  geom_point(size = 3, color = "#1F4E79") +
  geom_hline(yintercept = 5, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 10, linetype = "dotted", color = "grey50") +
  coord_flip() +
  labs(
    x = NULL,
    y = "Variance Inflation Factor (VIF)",
    title = "Assessment of multicollinearity (SAF model)",
    subtitle = "All VIF values are below conventional thresholds"
  ) +
  theme_minimal(base_family = "Times New Roman")



#predict comparable models ----
# Poisson ----
pois_RSF <- glm(
  hospital_attacks_RSF ~
    conflict_intensity +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population,
  data = panel_master_lag_actors,
  family = poisson()
)

summary(pois_RSF)
#deviance check, not good measure: deviance-based dispersion breaks down with extreme sparsity
deviance(pois_RSF) / df.residual(pois_RSF)
#so pearson-based dispersion
sum(residuals(pois_RSF, type = "pearson")^2) / df.residual(pois_RSF)
#3.016

pois_SAF <- glm(
  hospital_attacks_SAF ~
    conflict_intensity +
    territorial_losses_SAF +
    log_population,
  data = panel_master_lag_actors,
  family = poisson()
)

summary(pois_RSF)


#check overdisperion
performance::check_overdispersion(pois_RSF)
performance::check_overdispersion(pois_SAF)

# Negative Binomial----
nb_RSF <- glm.nb(
  hospital_attacks_RSF ~
    conflict_intensity +
    territorial_losses_RSF +
    territorial_losses_RSF_lag1 +
    log_population,
  data = panel_master_lag_actors
)

summary(nb_RSF)

nb_SAF <- glm.nb(
  hospital_attacks_SAF ~
    conflict_intensity +
    territorial_losses_SAF +
    log_population,
  data = panel_master_lag_actors
)

summary(nb_SAF)

#compare AIC (possion, negative binomial, ZINB)
AIC(pois_RSF, nb_RSF, zinb_RSF_t)
AIC(pois_SAF, nb_SAF, zinb_SAF_t)

#test overdispersion ----
lrtest(pois_RSF, nb_RSF)
lrtest(pois_SAF, nb_SAF)
#vuong test 
vuong(nb_RSF, zinb_RSF_t)
vuong(nb_SAF, zinb_SAF_t)
#check overdispersion ----
zinb_RSF_t$theta
zinb_SAF_t$theta
# - overdispersion is a thing - ZINB is valid!

#check residuals ----
plot_residuals_zinb_RSF <- plot(residuals(zinb_RSF_t, type = "pearson"))

print(plot_residuals_zinb_RSF)

plot_resdiuals_zinb_SAF <- plot(residuals(zinb_SAF_t, type = "pearson"))

print(plot_resdiuals_zinb_SAF)

#sparsity checks ----
#for territorial losses SAF lagged 
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(territorial_losses_SAF_lag1 != 0, na.rm = TRUE),
    share_nonzero = mean(territorial_losses_SAF_lag1 != 0, na.rm = TRUE)
  )
#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(territorial_losses_SAF_lag1 > 0,
           hospital_attacks_SAF > 0))

#territorial losses SAF 
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(territorial_losses_SAF != 0, na.rm = TRUE),
    share_nonzero = mean(territorial_losses_SAF != 0, na.rm = TRUE)
  )

#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(territorial_losses_SAF > 0,
           hospital_attacks_SAF > 0))

#territorial losses RSF lagged
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(territorial_losses_RSF_lag1 != 0, na.rm = TRUE),
    share_nonzero = mean(territorial_losses_RSF_lag1 != 0, na.rm = TRUE)
  )
#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(territorial_losses_RSF_lag1 > 0,
           hospital_attacks_RSF > 0))

#territorial losses RSF 
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(territorial_losses_RSF != 0, na.rm = TRUE),
    share_nonzero = mean(territorial_losses_RSF != 0, na.rm = TRUE)
  )
#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(territorial_losses_RSF > 0,
           hospital_attacks_RSF > 0))

#conflict Intensity 
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(conflict_intensity != 0, na.rm = TRUE),
    share_nonzero = mean(conflict_intensity != 0, na.rm = TRUE)
  )
#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(conflict_intensity > 0,
           hospital_attacks_RSF > 0))

#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(conflict_intensity > 0,
           hospital_attacks_SAF > 0))


#conflict intensity lagged
panel_master_lag_actors %>%
  summarise(
    n = n(),
    n_nonzero = sum(conflict_intensity_lag1 != 0, na.rm = TRUE),
    share_nonzero = mean(conflict_intensity_lag1 != 0, na.rm = TRUE)
  )
#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(conflict_intensity_lag1 > 0,
           hospital_attacks_RSF > 0))

#variable ever switch on when attacks occur?
with(panel_master_lag_actors,
     table(conflict_intensity_lag1 > 0,
           hospital_attacks_SAF > 0))


#separation check ----
with(panel_master_lag_actors,
     table(territorial_losses_SAF_lag1 > 0,
           hospital_attacks_SAF == 0))

#alternative check to sparsity check----
panel_master_lag_actors %>%
  filter(hospital_attacks_SAF > 0) %>%
  summarise(
    n_attacks = n(),
    share_loss_now = mean(territorial_losses_SAF != 0),
    share_loss_lag = mean(territorial_losses_SAF_lag1 != 0)
  )
#contingency table
with(panel_master_lag_actors,
     table(
       loss = territorial_losses_SAF_lag1 > 0,
       attack = hospital_attacks_SAF > 0
     ))

panel_master_lag_actors %>%
  filter(hospital_attacks_RSF > 0) %>%
  summarise(
    n_attacks = n(),
    share_loss_now = mean(territorial_losses_RSF != 0),
    share_loss_lag = mean(territorial_losses_RSF_lag1 != 0)
  )
```


